---
title: "CodeIndividualBasedSimulation"
author: "Yimen"
date: "8/30/2021"
output: html_document
---

The code below has the functions for the individual based simulations. Everything is stored in lists so it is easier to do use the different cores in the computer.
```{r packages}
require(arm)
require(dplyr)
require(MASS)
library(parallel)
library(ggplot2)
require(ggpubr)
require(RColorBrewer)

cols=brewer.pal(n = 3, name = "Set2")
alpha=0.2
```

Below is the main function to simulate the data.
```{r FunctionSimulate}
rm(list=ls())

sim2<-function(x){
  # x is a list with all the parameters to simulate
  #Generate different lists to store results
   dAll<-list(IndData=list(), PopData=list())
   d<-list(Ind=list(), Pop=list())
   dyears<-list()
   
   pop<-data.frame(sim=x$sim, year=1:x$n.years, N0=NA, N1=NA, N=NA, S=NA, z=NA)
    
   #Simulate among year variation. This creates the average survival and reproduction for each year.
   ##VYs is among year variation in survival
   ##VYr is among year variation in reproduction
   ##VYrs is the among-year covariance between survival and reproduction
    MY<-matrix(c(x$VYs, x$VYrs, x$VYrs, x$VYr),2,2)
    Ys<-mvrnorm(x$n.years, c(0,0), MY)
    Y<-data.frame(year=1:x$n.years, Yks=Ys[,1], Yrk=Ys[,2])
    
    # Simulate individual values for the first generation.
    ## Adult individuals
    ## N1 is the number of aldults in the founder population
    ## VIr is the among individual variance in number of recruits
    ## VIrs is the among individual variance in survival
    ## VIrs is a pheniotype that could affect reproduction or survival
    dA1<-data.frame(ID=1:x$N1, Iri=rnorm(x$N1, 0, x$VIr), Isi=rnorm(x$N1, 0, x$VIs), z=rnorm(x$N1, 0, sqrt(x$VIrs)))
    dA1$age<-1
    
    ##First breeding year individuals are simulated in the same way
    dJ1<-data.frame(ID=(x$N1+1):(x$N0+x$N1), Iri=rnorm(x$N0, 0, x$VIr), Isi=rnorm(x$N0, 0, x$VIs), z=rnorm(x$N0, 0, sqrt(x$VIrs)))
    dJ1$age<-0
    
    ###Combine the adult and juvenile data
    d1<-rbind(dA1, dJ1)
    
    ###Average phenotype in the founder population
    d1$z_bar<-mean(d1$z)
    
    ###Adds the year means for all individuals in the same year
    d1$year<-1
    d1$Yks<-Y$Yks[1] 
    d1$Yrk<- Y$Yrk[1]
    
    ###Adds the starting population sizes of juveniles and adults
    d1$N0<-x$N0
    d1$N1<-x$N1
    d1$N<-(x$N1+x$N0)
    
    ###Generate residual values for each year (i.e) stocahsiticity affecting an individuals reproduction or survival in a given year
    MR<-matrix(c(x$Vss, x$Vsrs, x$Vsrs, x$Vsr),2,2)
    Es<-mvrnorm(nrow(d1), c(0,0), MR)
     
    ##Model survival
    d1$sm <- x$s + d1$Isi + x$Bzs*(d1$z) + d1$Yks  + d1$N*x$Gs + Es[,1]
    d1$pr = (1/(1+exp(-d1$sm))) ### transform to individual survival probabilities
    d1$surv<-rbinom(length(d1$sm),1,d1$pr) ###Realization for survival
    
    ###Model reproduction
    d1$rm<- x$r + d1$Iri + x$Bzr*(d1$z) + d1$Yrk  + d1$N*x$Gr +  Es[,2]
    d1$recruits<-rpois(length(d1$rm), exp(d1$rm)) ##The exponent of the latent trait defines the mean     for the poisson process
    
    ### Add reproduction and survival to estimate yearly individual fitness
    d1$w<-d1$recruits + d1$surv 
    d1$rw<-d1$w/mean(d1$w) ####Relative fitness
    
    ##Create a dataframe with population level parameters
    pop$sim[1]<-x$sim
    pop$N0[1]<-x$N0
    pop$N1[1]<-x$N1
    pop$N[1]<-x$N0+x$N1
    pop$lambda[1]<-sum(d1$recruits + d1$surv)/(N0+N1)
    pop$z[1]<-d1$z_bar[1]
    pop$deltaz[1]<-d1$z_bar[1]+cov(d1$z, d1$rw)

    ##Store the individual data in a list 
    dyears[[1]]<-d1 
    
    #Other generations (year). 
    ##The tryCatch function prevents the loop to abort if there is and error.
    tryCatch({
      for(i in 2:x$n.years){
    
    ##Create temporary data with individuals of the previous gen.
        dtmp<-as.data.frame(dyears[[i-1]])
  
     ###Survival adults
        dA2<-dtmp[dtmp$surv==1, c("ID", "Iri", "Isi", "z")]
        dA2$age<-1 + dtmp$age[dtmp$surv==1]
    
    ###Creat the new recruting indivdiuals. They have the same characteristics of the parents         
        Iri=rep(dtmp$Iri, dtmp$recruits)
        Isi=rep(dtmp$Isi, dtmp$recruits)
        z=rep(dtmp$z, dtmp$recruits)
    
    ####If there are new recruits it creates there phneotypes and intrinsic survival and reproduction         merits.      
          if (length(Iri)>0) {
          dJ2<-data.frame(ID=paste(i, rep(dtmp$ID, dtmp$recruits), 1:sum(dtmp$recruits)))
          dJ2$Iri=Iri*x$h2r + rnorm(length(Iri), 0, sqrt(x$VIr-x$h2r^2))
          dJ2$Isi=Isi*x$h2s + rnorm(length(Isi), 0, sqrt(x$VIs-x$h2s^2))
          dJ2$z= z + rnorm(length(z), 0, sqrt(0.5))
          dJ2$age<-0
          }
          
    ### If some individuals survive and some reproduced combines the data into d2
         if(nrow(dA2)>0 & length(Iri)>0){
            d2<-rbind(dA2,  dJ2)
          }
    
    ### If some individuals survive and but nobody reproduced d2 is composed by only adults of previous year 
         if(nrow(dA2)>0 & length(Iri)==0){
            d2<-dA2
          }
  ### No adult survives but some reprodcue, d2 is composed of only juveniles.       
         if(nrow(dA2)==0 & length(Iri)>0){
            d2<-dJ2
          }

  ### If there is no reproduction and no one survives end the loop          
        if(nrow(dA2)==0 & length(Iri)==0){
            break
          }
        
        d2$year<-i  ##Add year to the individual data
        d2$Yks<-Y$Yks[i] ##Add average year survival to the individual data
        d2$Yrk<- Y$Yrk[i] ##Add average year reproduction to the individual data
       
        d2$N0<-sum(dtmp$recruits) ##Number of juveniles
        d2$N1<-sum(dtmp$surv) ##Number of adults
        d2$N<-sum(dtmp$recruits + dtmp$surv) ##Population size 
        
        ###Stocastic variation for individuals yearly survival and reproduction
        MR<-matrix(c(x$Vss, x$Vsrs, x$Vsrs, x$Vsr),2,2)
        Es<-mvrnorm(nrow(d2), c(0,0), MR)
        
        ##Age as a categorical value and mean phenotype
        age2<-d2$age
        age2[d2$age>0]<-1
        d2$z_bar<-mean(c(dA2$z,dJ2$z))
        
        ##Same as for the founder population but with extra terms reflecting the interactions between population size and mean phenotype
        d2$sm <- x$s + x$as*age2  + d2$Isi + d2$Yks + x$Bzs*d2$z +  d2$N*x$Gs +  x$BZs*d2$z_bar + x$BZns*d2$N*d2$z_bar + x$Bzns*d2$N*d2$z  + x$BZzs*d2$z*d2$z_bar+ x$BZzns*d2$z*d2$z_bar*d2$N + Es[,1]
        
        d2$pr = (1/(1+exp(-d2$sm)))  
        d2$surv<-rbinom(length(d2$sm),1,d2$pr) 
        
        d2$rm<-x$r + x$ar*age2 +  d2$Iri + d2$Yrk + x$Bzr*d2$z + x$Bzr2*d2$z^2 +  d2$N*x$Gr +  x$BZr*d2$z_bar + x$BZnr*d2$N*d2$z_bar + x$Bznr*d2$N*d2$z  + x$BZzr*d2$z*d2$z_bar+ x$BZznr*d2$z*d2$z_bar*d2$N + Es[,2]  
        d2$recruits<-rpois(length(d2$rm), exp(d2$rm))
       
        d2$w<-d2$recruits + d2$surv
        d2$rw<-d2$w/mean(d2$w)
        
        dyears[[i]]<-d2
        
        pop$sim[i]<-x$sim ##Simulation ID
        pop$N0[i]<-sum(dtmp$recruits)
        pop$N1[i]<-sum(dtmp$surv)
        pop$N[i]<- sum(dtmp$recruits + dtmp$surv)
        pop$z[i]<-d2$z_bar[1]
        pop$lambda[i]<- sum(d2$recruits + d2$surv)/pop$N[i]
        pop$deltaz[i]<-mean(d2$z)+cov(d2$z, d2$rw) ##Expected change in mean phenotype
              }
    }, error=function(e){})
  
    ##Combine individual data for all years
    dtmp1<-do.call(rbind.data.frame, dyears)
    dtmp1$sim<-x$sim ##Simulation ID
    dAll$IndData<-dtmp1 ##Get it into a list
    dAll$PopData<-pop ##Get population data into a list
    dAll
    }


##Creates a list of parameter values for the simulations to use lapply to generate all simulations
sim.pops<-function(n.years=10, n.sims=1, s, r, VYs=0, VYr=0, VYrs=0, VIs=0, VIr=0, VIrs=0, h2r=0, h2s=0, h2rs=0, Vss=0, Vsr=0, Vsrs=0, Gs=0, Gr=0, N0=10, N1=10, ar=0, as=0, Bzr=0, Bzr2=0, Bzs=0, BZr=0, BZs=0, BZnr=0, BZns=0, Bznr=0, Bzns=0, BZzr=0,  BZzs=0, BZznr=0,  BZzns=0 )
{  
lists<-list()
d<-list()

for(j in 1:n.sims){
lists[[j]]<-data.frame(n.years=n.years, s=s, r=r, VYs=VYs, VYr=VYr, VYrs=VYrs, VIs=VYrs, VIr=VIr, VIrs=VIrs, h2r=h2r, h2s=h2s, h2rs=h2rs, Vss=Vss, Vsr=Vsr, Vsrs=Vsrs, Gs=Gs, Gr=Gr, N0=N0, N1=N1, ar=ar, as=as, Bzr=Bzr, Bzr2=Bzr2, Bzs=Bzs, BZr=BZr, BZs=BZs, BZnr=BZnr, BZns=BZns, Bznr=Bznr, Bzns=Bzns, BZzr=BZzr,  BZzs=BZzs, BZznr=BZznr,  BZzns=BZzns, sim=j) 
}    
 
res<-mclapply(lists, sim2, mc.cores=4)
d$Ind<-do.call(rbind.data.frame, lapply(res, '[[', 1))
d$Pop<-do.call(rbind.data.frame, lapply(res, '[[', 2))
d$Params<-as.data.frame(lists[1])
d
}




```

Below is a bunch of functions to analyze the data. We can look at this later.
```{r FunctionAnalyze}
Nbar<-function(x,y,z){
mp1<-tapply(x$Pop$N, x$Pop$year, mean, na.rm=TRUE)
mpd1<-data.frame(year=1:length(mp1), N=mp1, scen="1")

mp2<-tapply(y$Pop$N, y$Pop$year, mean, na.rm=TRUE) 
mpd2<-data.frame(year=1:length(mp2), N=mp2, scen="2")

mp3<-tapply(z$Pop$N, z$Pop$year, mean, na.rm=TRUE)
mpd3<-data.frame(year=1:length(mp3), N=mp3, scen="3")

mpd<-rbind(mpd1,mpd2,mpd3)
mpd
}

zbar<-function(x,y,z){
mp1<-tapply(x$Pop$z, x$Pop$year, mean, na.rm=TRUE)
mpd1<-data.frame(year=1:length(mp1), z=mp1, scen="1")

mp2<-tapply(y$Pop$z, y$Pop$year, mean, na.rm=TRUE) 
mpd2<-data.frame(year=1:length(mp2), z=mp2, scen="2")

mp3<-tapply(z$Pop$z, z$Pop$year, mean, na.rm=TRUE)
mpd3<-data.frame(year=1:length(mp3), z=mp3, scen="3")

mpd<-rbind(mpd1,mpd2,mpd3)
mpd
}

Ext<-function(x,y,z){
x$Pop$Ext<-is.na(x$Pop$N)
x$Pop$Ext<-as.numeric(x$Pop$Ext)
Ext<-tapply(x$Pop$Ext, x$Pop$year, sum)/n.sims 
med1<-data.frame(year=1:length(Ext), e=Ext, scen="1")

y$Pop$Ext<-is.na(y$Pop$N)
y$Pop$Ext<-as.numeric(y$Pop$Ext)
Ext2<-tapply(y$Pop$Ext, y$Pop$year, sum)/n.sims 
med2<-data.frame(year=1:length(Ext2), e=Ext2, scen="2")

z$Pop$Ext<-is.na(z$Pop$N)
z$Pop$Ext<-as.numeric(z$Pop$Ext)
Ext3<-tapply(z$Pop$Ext, z$Pop$year, sum)/n.sims 
med3<-data.frame(year=1:length(Ext3), e=Ext3, scen="3")

med<-rbind(med1,med2,med3)
med
}



model.1<-function(x){
mod<-lm(w~N, data=x)
B1=coef(mod)[1]
Bn=coef(mod)[2]

EN=as.vector(-(B1-1)/Bn)

res<-data.frame(N=x$N[x$year==100][1], z=mean(x$z[x$year==100]), EN=EN, Ez=NA)
res$biasN=as.vector(res$EN-res$N)
res
}



model.2<-function(x){
x$z2<-x$z^2
mod<-lm(w~N + z + z2, data=x)
B1=coef(mod)[1]
Bn=coef(mod)[2]
Bl=coef(mod)[3]
Bq=coef(mod)[4]
zp=(-Bl/(2*Bq))
EN=as.vector(-(B1 + Bl*zp + Bq*zp^2+Bq*var(x$z[x$year==100])-1)/Bn)

res<-data.frame(N=x$N[x$year==100][1], z=mean(x$z[x$year==100]), EN=EN, Ez=zp)
res$biasN=as.vector(res$EN-res$N)
res
}

model.3<-function(x){
x$z2<-x$z^2
mod<-lm(w~N + z + z2 +  z_bar , data=x)
B1=coef(mod)[1]
Bn=coef(mod)[2]
Bl=coef(mod)[3]
Bq=coef(mod)[4]
Bz=coef(mod)[5]
zp=(-Bl/(2*Bq))

EN=(-1 + B1 + Bl*zp + Bz*zp+  Bq*zp^2  + Bq*var(x$z[x$year==100]))/-(Bn)


res<-data.frame(N=x$N[x$year==100][1], z=mean(x$z[x$year==100]), EN=EN, Ez=zp)
res$biasN=as.vector(res$EN-res$N)
res
}


model.4<-function(x){
x$z2<-x$z^2
mod<-lm(w~ N + z + z2 + z_bar + z_bar:N , data=x)
B1=coef(mod)[1]
Bn=coef(mod)[2]
Bl=coef(mod)[3]
Bq=coef(mod)[4]
Bz=coef(mod)[5]
Bnz=coef(mod)[6]
zp=(-Bl/(2*Bq))

EN=(-1 + B1 + Bl*zp + Bz*zp+  Bq*zp^2  + Bq*var(x$z[x$year==100]))/-(Bn + Bnz*zp)


res<-data.frame(N=x$N[x$year==100][1], z=mean(x$z[x$year==100]), EN=EN, Ez=zp)
res$biasN=as.vector(res$EN-res$N)
res
}


model.5<-function(x){
x$z2<-x$z^2
mod<-lm(w~ N + z + z2 + z:N  , data=x)
B0=coef(mod)[1]
Bn=coef(mod)[2]
Bl=coef(mod)[3]
Bq=coef(mod)[4]
Bzn=coef(mod)[5]
N=x$N[x$year==100][1]
s=var(x$z[x$year==100])

zp<--(Bl - (Bl*Bzn - 2*Bn*Bq + 2*sqrt(Bq^2*Bzn^2*s + B0*Bq*Bzn^2 - Bl*Bn*Bq*Bzn + Bn^2*Bq^2 - Bq*Bzn^2))/Bzn)/(2*Bq)

EN=-(Bl*Bzn - 2*Bn*Bq + 2*sqrt(Bq^2*Bzn^2*s + B0*Bq*Bzn^2 - Bl*Bn*Bq*Bzn + Bn^2*Bq^2 - Bq*Bzn^2))/Bzn^2
res<-data.frame(EN=EN, N=N, Ez=zp, z=mean(x$z[x$year==100]))
res$biasN=as.vector(res$EN-res$N)
res
}


model.6<-function(x){
x$z2<-x$z^2
mod<-lm(w~ N + z + z2 + z_bar + z:z_bar  , data=x)
B0=coef(mod)[1]
Bn=coef(mod)[2]
Bl=coef(mod)[3]
Bq=coef(mod)[4]
Bz=coef(mod)[5]
Bzz=coef(mod)[6]
zp<--(Bl)/(2*Bq + Bzz)

EN=(-1 + B0 + (Bl + Bz)*zp  +  (Bzz + Bq)*zp^2 + Bq*var(x$z[x$year==100]))/-Bn

res<-data.frame(N=x$N[x$year==100][1], z=mean(x$z[x$year==100]), EN=EN, Ez=zp)
res$biasN=as.vector(res$EN-res$N)
res
}


model.7<-function(x){
x$z2<-x$z^2
mod<-lm(w~ N + z + z2 + z_bar + z_bar:N + z:N + z_bar:z + z_bar:N:z, data=x)
B0=coef(mod)[1]
Bn=coef(mod)[2]
Bl=coef(mod)[3]
Bq=coef(mod)[4]
Bz=coef(mod)[5]
BNZ=coef(mod)[6]
BNz=coef(mod)[7]
BZz=coef(mod)[8]
BNZz=coef(mod)[9]

N=x$N[x$year==100][1]

zp<--(Bl + BNz*N)/(2*Bq + BZz + BNZz*N)

EN=(-1 + B0 + (Bl + Bz)*zp  +  (BZz + Bq)*zp^2 + Bq*var(x$z[x$year==100]))/-(Bn + BNz*zp + BNZz*zp^2 + BNZ*zp) 

res<-data.frame(EN=EN, N=N, Ez=zp, z=mean(x$z[x$year==100]))
res$biasN=as.vector(res$EN-res$N)
res
}



bias<-function(x,y,z, model){
ld3<-ld2<-ld1<-list()
 
for(i in 1:n.sims){
ld1[[i]]<-as.data.frame(x$Ind[x$Ind$sim==i,])
ld2[[i]]<-as.data.frame(y$Ind[y$Ind$sim==i,])
ld3[[i]]<-as.data.frame(z$Ind[z$Ind$sim==i,])
}

an1<-do.call(rbind.data.frame,mclapply(ld1, model, mc.cores=3))
an2<-do.call(rbind.data.frame,mclapply(ld2, model, mc.cores=3))
an3<-do.call(rbind.data.frame,mclapply(ld3, model, mc.cores=3))
b<-rbind(apply(an1,2,mean, na.rm=TRUE), apply(an2,2,mean, na.rm=TRUE), apply(an3,2,mean, na.rm=TRUE))
}

param<-function(x,y,z){
p<-c("Gr", "Bzr", "Bzr2", "BZr", "BZnr", "BZzr", "Bznr", "BZznr")  
rbind(x$Params[,p], y$Params[,p], z$Params[,p])
}

```

The first scenario of the ms I sent.
##Density regulation
```{r SelectionSimulation, cache=TRUE}
n.years<-100
n.sims<-100
N0<-5
N1<-5
s<--0.1
r<-0.1
Gr<--0.02
Gs<--0.000

res1<-sim.pops(n.years=n.years, n.sims=n.sims, s=s, r=r, VIrs=1, h2rs=0.99, Gs=Gs, Gr=Gr, N0=N0, N1=N1, Bzr=0,  Bzr2 =0, Bzs=0, BZr=0, BZs=0, BZnr=0, BZns=0, Bznr=0,  Bzns=0, BZzr=0,  BZzs=0, BZznr=0,  BZzns=0)

res2<-sim.pops(n.years=n.years, n.sims=n.sims, s=s, r=r, VIrs=1, h2rs=0.99, Gs=Gs, Gr=Gr, N0=N0, N1=N1, Bzr=0.2,  Bzr2 =-0.02, Bzs=0, BZr=0, BZs=0, BZnr=0, BZns=0, Bznr=0,  Bzns=0, BZzr=0,  BZzs=0, BZznr=0,  BZzns=0 )

res3<-sim.pops(n.years=n.years, n.sims=n.sims, s=s, r=r, VIrs=1, h2rs=0.99, Gs=Gs, Gr=Gr, N0=N0, N1=N1, Bzr=-0.2,  Bzr2 =-0.02, Bzs=0, BZr=0, BZs=0, BZnr=0, BZns=0, Bznr=0,  Bzns=0, BZzr=0,  BZzs=0, BZznr=0,  BZzns=0 )

mpd2<-Nbar(res1,res2,res3)
mzd2<-zbar(res1,res2,res3)
med2<-Ext(res1,res2,res3)
bias2<-bias(res1,res2,res3, model.2)
par2<-param(res1, res2, res3)

BiasPar2<-cbind(bias2,par2)

```

```{r SelectionFigure}
p3<-ggplot(mpd2, aes(x = year, y = N, color=scen)) + geom_line( show.legend = FALSE)+ scale_color_manual(values=cols)   + geom_line(data = res1$Pop, alpha = alpha, col=cols[1])  + geom_line(data = res2$Pop, alpha = alpha, col=cols[2]) + geom_line(data = res3$Pop, alpha = alpha, col=cols[3]) + theme_classic()  + labs(title="B)", x="Year", y = "N") + ylim(0,120) + theme(legend.text.align = 0)

p4<-ggplot(mzd2, aes(x = year, y = z, color=scen)) + geom_line(alpha = .8) + scale_color_manual(values=cols, 
                       name="Parameters",
                       breaks=c("1", "2", "3"),
labels=c(expression(paste(beta[n],"=-0.02,  ",   beta[z],"=0")),expression(paste(beta[n],"=-0.2, ",   beta[z],"=0.02 ")),expression(paste(beta[n],"=-0.02,",   beta[z],"=-0.02"))))  + geom_line(data = res1$Pop, alpha = alpha, col=cols[1])  + geom_line(data = res2$Pop, alpha = alpha, col=cols[2]) + geom_line(data = res3$Pop, alpha = alpha, col=cols[3]) + theme_classic()  + labs(title="", x="Year", y = "Mean phenotype (z)") + ylim(-10,20) + theme(legend.text.align = 0)

ggarrange( p3,  p4, ncol=2, nrow=1, widths=c(1,1.4)) 

```

We can check step by step what the code to simulate the data does by creating a data frame with parameter values we need.
```{r}

n.years<-100
n.sims<-100
N0<-5
N1<-5
s<--0.1
r<-0.1
Gr<--0.01
Gs<--0.000

x<-data.frame(n.years=n.years, s=s, r=r, VYs=0, VYr=0, VYrs=0, VIs=1, VIr=1, VIrs=1, h2r=1, h2s=1, h2rs=1, Vss=0, Vsr=0, Vsrs=0, Gs=Gs, Gr=Gr, N0=N0, N1=N1, ar=0, as=0, Bzr=0, Bzr2=0, Bzs=0, BZr=0, BZs=0, BZnr=0, BZns=0, Bznr=0, Bzns=0, BZzr=0,  BZzs=0, BZznr=0,  BZzns=0, sim=1) 


  # x is a list with all the parameters to simulate
  #Generate different lists to store results
   dAll<-list(IndData=list(), PopData=list())
   d<-list(Ind=list(), Pop=list())
   dyears<-list()
   
   pop<-data.frame(sim=x$sim, year=1:x$n.years, N0=NA, N1=NA, N=NA, S=NA, z=NA)
    
   #Simulate among year variation. This creates the average survival and reproduction for each year.
   ##VYs is among year variation in survival
   ##VYr is among year variation in reproduction
   ##VYrs is the among-year covariance between survival and reproduction
    MY<-matrix(c(x$VYs, x$VYrs, x$VYrs, x$VYr),2,2)
    Ys<-mvrnorm(x$n.years, c(0,0), MY)
    Y<-data.frame(year=1:x$n.years, Yks=Ys[,1], Yrk=Ys[,2])
    
    # Simulate individual values for the first generation.
    ## Adult individuals
    ## N1 is the number of aldults in the founder population
    ## VIr is the among individual variance in number of recruits
    ## VIrs is the among individual variance in survival
    ## VIrs is a pheniotype that could affect reproduction or survival
    dA1<-data.frame(ID=1:x$N1, Iri=rnorm(x$N1, 0, x$VIr), Isi=rnorm(x$N1, 0, x$VIs), z=rnorm(x$N1, 0, sqrt(x$VIrs)))
    dA1$age<-1
    
    ##First breeding year individuals are simulated in the same way
    dJ1<-data.frame(ID=(x$N1+1):(x$N0+x$N1), Iri=rnorm(x$N0, 0, x$VIr), Isi=rnorm(x$N0, 0, x$VIs), z=rnorm(x$N0, 0, sqrt(x$VIrs)))
    dJ1$age<-0
    
    ###Combine the adult and juvenile data
    d1<-rbind(dA1, dJ1)
    
    ###Average phenotype in the founder population
    d1$z_bar<-mean(d1$z)
    
    ###Adds the year means for all individuals in the same year
    d1$year<-1
    d1$Yks<-Y$Yks[1] 
    d1$Yrk<- Y$Yrk[1]
    
    ###Adds the starting population sizes of juveniles and adults
    d1$N0<-x$N0
    d1$N1<-x$N1
    d1$N<-(x$N1+x$N0)
    
    ###Generate residual values for each year (i.e) stocahsiticity affecting an individuals reproduction or survival in a given year
    MR<-matrix(c(x$Vss, x$Vsrs, x$Vsrs, x$Vsr),2,2)
    Es<-mvrnorm(nrow(d1), c(0,0), MR)
     
    ##Model survival
    d1$sm <- x$s + d1$Isi + x$Bzs*(d1$z) + d1$Yks  + d1$N*x$Gs + Es[,1]
    d1$pr = (1/(1+exp(-d1$sm))) ### transform to individual survival probabilities
    d1$surv<-rbinom(length(d1$sm),1,d1$pr) ###Realization for survival
    
    ###Model reproduction
    d1$rm<- x$r + d1$Iri + x$Bzr*(d1$z) + d1$Yrk  + d1$N*x$Gr +  Es[,2]
    d1$recruits<-rpois(length(d1$rm), exp(d1$rm)) ##The exponent of the latent trait defines the mean     for the poisson process
    
    ### Add reproduction and survival to estimate yearly individual fitness
    d1$w<-d1$recruits + d1$surv 
    d1$rw<-d1$w/mean(d1$w) ####Relative fitness
    
    ##Create a dataframe with population level parameters
    pop$sim[1]<-x$sim
    pop$N0[1]<-x$N0
    pop$N1[1]<-x$N1
    pop$N[1]<-x$N0+x$N1
    pop$lambda[1]<-sum(d1$recruits + d1$surv)/(N0+N1)
    pop$z[1]<-d1$z_bar[1]
    pop$deltaz[1]<-d1$z_bar[1]+cov(d1$z, d1$rw)

    ##Store the individual data in a list 
    dyears[[1]]<-d1 
    
    #Other generations (year). 
    ##The tryCatch function prevents the loop to abort if there is and error.
    tryCatch({
      for(i in 2:x$n.years){
    i=2
    ##Create temporary data with individuals of the previous gen.
        dtmp<-as.data.frame(dyears[[i-1]])
  
     ###Survival adults
        dA2<-dtmp[dtmp$surv==1, c("ID", "Iri", "Isi", "z")]
        dA2$age<-1 + dtmp$age[dtmp$surv==1]
    
    ###Creat the new recruting indivdiuals. They have the same characteristics of the parents         
        Iri=rep(dtmp$Iri, dtmp$recruits)
        Isi=rep(dtmp$Isi, dtmp$recruits)
        z=rep(dtmp$z, dtmp$recruits)
    
    ####If there are new recruits it creates there phneotypes and intrinsic survival and reproduction         merits.      
          if (length(Iri)>0) {
          dJ2<-data.frame(ID=paste(i, rep(dtmp$ID, dtmp$recruits), 1:sum(dtmp$recruits)))
          dJ2$Iri=Iri*x$h2r + rnorm(length(Iri), 0, sqrt(x$VIr-x$h2r^2))
          dJ2$Isi=Isi*x$h2s + rnorm(length(Isi), 0, sqrt(x$VIs-x$h2s^2))
          dJ2$z= z + rnorm(length(z), 0, sqrt(0.5))
          dJ2$age<-0
          }
          
    ### If some individuals survive and some reproduced combines the data into d2
         if(nrow(dA2)>0 & length(Iri)>0){
            d2<-rbind(dA2,  dJ2)
          }
    
    ### If some individuals survive and but nobody reproduced d2 is composed by only adults of previous year 
         if(nrow(dA2)>0 & length(Iri)==0){
            d2<-dA2
          }
  ### No adult survives but some reprodcue, d2 is composed of only juveniles.       
         if(nrow(dA2)==0 & length(Iri)>0){
            d2<-dJ2
          }

  ### If there is no reproduction and no one survives end the loop          
        if(nrow(dA2)==0 & length(Iri)==0){
            break
          }
        
        d2$year<-i  ##Add year to the individual data
        d2$Yks<-Y$Yks[i] ##Add average year survival to the individual data
        d2$Yrk<- Y$Yrk[i] ##Add average year reproduction to the individual data
       
        d2$N0<-sum(dtmp$recruits) ##Number of juveniles
        d2$N1<-sum(dtmp$surv) ##Number of adults
        d2$N<-sum(dtmp$recruits + dtmp$surv) ##Population size 
        
        ###Stocastic variation for individuals yearly survival and reproduction
        MR<-matrix(c(x$Vss, x$Vsrs, x$Vsrs, x$Vsr),2,2)
        Es<-mvrnorm(nrow(d2), c(0,0), MR)
        
        ##Age as a categorical value and mean phenotype
        age2<-d2$age
        age2[d2$age>0]<-1
        d2$z_bar<-mean(c(dA2$z,dJ2$z))
        
        ##Same as for the founder population but with extra terms reflecting the interactions between population size and mean phenotype
        d2$sm <- x$s + x$as*age2  + d2$Isi + d2$Yks + x$Bzs*d2$z +  d2$N*x$Gs +  x$BZs*d2$z_bar + x$BZns*d2$N*d2$z_bar + x$Bzns*d2$N*d2$z  + x$BZzs*d2$z*d2$z_bar+ x$BZzns*d2$z*d2$z_bar*d2$N + Es[,1]
        
        d2$pr = (1/(1+exp(-d2$sm)))  
        d2$surv<-rbinom(length(d2$sm),1,d2$pr) 
        
        d2$rm<-x$r + x$ar*age2 +  d2$Iri + d2$Yrk + x$Bzr*d2$z + x$Bzr2*d2$z^2 +  d2$N*x$Gr +  x$BZr*d2$z_bar + x$BZnr*d2$N*d2$z_bar + x$Bznr*d2$N*d2$z  + x$BZzr*d2$z*d2$z_bar+ x$BZznr*d2$z*d2$z_bar*d2$N + Es[,2]  
        d2$recruits<-rpois(length(d2$rm), exp(d2$rm))
       
        d2$w<-d2$recruits + d2$surv
        d2$rw<-d2$w/mean(d2$w)
        
        dyears[[i]]<-d2
        
        pop$sim[i]<-x$sim ##Simulation ID
        pop$N0[i]<-sum(dtmp$recruits)
        pop$N1[i]<-sum(dtmp$surv)
        pop$N[i]<- sum(dtmp$recruits + dtmp$surv)
        pop$lambda[i]<- sum(d2$recruits + d2$surv)/pop$N[i]
        pop$z[i]<-d2$z_bar[1]
        pop$deltaz[i]<-mean(d2$z)+cov(d2$z, d2$rw) ##Expected change in mean phenotype
              }
    }, error=function(e){})
  
    ##Combine individual data for all years
    dtmp1<-do.call(rbind.data.frame, dyears)
    dtmp1$sim<-x$sim ##Simulation ID
    dAll$IndData<-dtmp1 ##Get it into a list
    dAll$PopData<-pop ##Get population data into a list
    dAll
    }


##Creates a list of parameter values for the simulations to use lapply to generate all simulations
sim.pops<-function(n.years=10, n.sims=1, s, r, VYs=0, VYr=0, VYrs=0, VIs=0, VIr=0, VIrs=0, h2r=0, h2s=0, h2rs=0, Vss=0, Vsr=0, Vsrs=0, Gs=0, Gr=0, N0=10, N1=10, ar=0, as=0, Bzr=0, Bzr2=0, Bzs=0, BZr=0, BZs=0, BZnr=0, BZns=0, Bznr=0, Bzns=0, BZzr=0,  BZzs=0, BZznr=0,  BZzns=0 )
{  
lists<-list()
d<-list()

for(j in 1:n.sims){
lists[[j]]<-data.frame(n.years=n.years, s=s, r=r, VYs=VYs, VYr=VYr, VYrs=VYrs, VIs=VYrs, VIr=VIr, VIrs=VIrs, h2r=h2r, h2s=h2s, h2rs=h2rs, Vss=Vss, Vsr=Vsr, Vsrs=Vsrs, Gs=Gs, Gr=Gr, N0=N0, N1=N1, ar=ar, as=as, Bzr=Bzr, Bzr2=Bzr2, Bzs=Bzs, BZr=BZr, BZs=BZs, BZnr=BZnr, BZns=BZns, Bznr=Bznr, Bzns=Bzns, BZzr=BZzr,  BZzs=BZzs, BZznr=BZznr,  BZzns=BZzns, sim=j) 
}    
 
res<-mclapply(lists, sim2, mc.cores=4)
d$Ind<-do.call(rbind.data.frame, lapply(res, '[[', 1))
d$Pop<-do.call(rbind.data.frame, lapply(res, '[[', 2))
d$Params<-as.data.frame(lists[1])
d
}

```

